 

### 2021-07-14 vertical-thermal-review.Rmd
Update vertical-thermal-review.Rmd

```{r eval=FALSE, echo=FALSE}
# refs in this table need to go in the nocite field below (easiest in alphabetical order)
```
---
nocite: |
@mau_temperate_2018; @cobleLightDrivesVertical2014; @sackHowStrongIntracanopy2006; @chinLeafAcclimationLight2017; @wykaResponsesLeafStructure2012; @athertonSpatialVariationLeaf2017; @kenzoHeightrelatedChangesLeaf2015; @kusiPlasticLeafMorphology2020; @dangProfilesPhotosyntheticallyActive1997;@gebauerEffectsProlongedDrought2015; @marencoLeafTraitPlasticity2017; @kafutiFoliarWoodTraits2020; @vanwittenbergheVariabilityStomatalConductance2012; @zhang_photosynthetic_2012; @weerasingheCanopyPositionAffects2014; @oldhamHydrostaticGradientNot2010; @ichieEcologicalDistributionLeaf2016; @gregoriouEffectsReducedIrradiance2007; @levizouNondestructiveAssessmentLeaf2005; @liakouraTrichomeDensityIts1997; @fauset_differences_2018; @niinemetsAcclimationHighIrradiance1998; Ishida et al. 1999; @millenLeafAngleAdaptive1979; @smithShootStructuralEffects1988; @hadleyInfluenceKrummholzMat1987; @panditharathnaChangeLeafStructure2008; @baltzerLeafOpticalResponses2005; @dietzVerticalPatternsDuration2007; @cobleHowVerticalPatterns2016a; @scartazzaInvestigatingEuropeanBeech2016; @duursmaVerticalCanopyGradients2006; @harleyEffectsLightTemperature1996; @hernandezSimilarTemperatureDependence2020; @turnbullScalingFoliarRespiration2003; @chenLeafEconomicsSpectrum2020; @vandewegPhotosyntheticParametersDark2012; @cavaleri_foliar_2008; @konigerXanthophyllCyclePigmentsPhotosynthetic1995;@matsubaraSunshadePatternsLeaf2009; Harris and Medina 2013; @hansenVariationPigmentComposition2002a; @poorterLeafOpticalProperties1995; @cobleHowVerticalPatterns2016; @niinemetsCanopyGradientsLeaf2004; @poorterLeafOpticalProperties2000
---

@@

Further, higher heat tolerance in exposed canopies is facilitated by isoprenes in species capable of emitting isoprenes [REF].

@@

```{r eval=FALSE, echo=FALSE}
# refs in this table need to go in the nocite field below (easiest in alphabetical order)
```
---
nocite: | @kafutiFoliarWoodTraits2020; @vanwittenbergheVariabilityStomatalConductance2012a; @roberts_stomatal_1990; @dangProfilesPhotosyntheticallyActive1997; @marencoLeafTraitPlasticity2017; @ambroseEffectsHeightTreetop2010; @zweifelMiddayStomatalClosure2002; @slot_photosynthetic_2019; @hernandezSimilarTemperatureDependence2020; @urbanInductionPhotosynthesisImportance2007;Carter and Cavaleri 2018; @martin_boundary_1999; @mau_temperate_2018; @kosugiSeasonalFluctuationsTemperature2006; @niinemetsVariabilityLeafMorphology2015; @bachofenLightVPDGradients2020; Hamerlynck and Knapp 1994; @cobleHowVerticalPatterns2016; @wykaResponsesLeafStructure2012; @rijkersEffectTreeHeight2000a; Ishida et al. 1999; @weerasingheCanopyPositionAffects2014; @scartazzaInvestigatingEuropeanBeech2016; @millerOnlySunlitLeaves2021; Harris and Medina 2013; @legnerWithincanopyVariationPhotosynthetic2014; @kitaoCanopyNitrogenDistribution2018; @fauset_differences_2018; @rey-sanchez_spatial_2016; @mullerEvidenceEfficientNonevaporative2021; @curtis_intracanopy_2019; @meirLeafRespirationTwo2001; @turnbullScalingFoliarRespiration2003; @arakiVerticalSeasonalVariations2017; @bolstad_foliar_1999; @kenzoHeightrelatedChangesLeaf2015; @harleyEffectsLightTemperature1996;@xuSeasonalVariationTemperature2006; @harleyEnvironmentalControlsIsoprene1997; @niinemetsHowLightTemperature2015; @sharkeyFutureIsopreneEmission2014b; @sharkeyFutureIsopreneEmission2014; @simpragaVerticalCanopyGradient2013; @athertonSpatialVariationLeaf2017
---

@@

**Vertical gradients affect tree metabolism and growth.**
Whole-tree transpiration and gross primary production (GPP) increase in tandem with increasing tree size, where size is most commonly measured diameter breast height (DBH), but functionally linked to metrics such as foliage biomass, crown volume, or height, all of which scale with DBH [**REFS**, @anderson-teixeiraSizeRelatedScaling2015].
Specifically, increases are linked increased leaf area and in part to increasing probability that the crown is in the canopy [@muller-landauTestingMetaabolicEcology2006], where higher light availability and associated elevation of $g_s$ and $A_A$ result in higher leaf area-specific metabolism (Table 2).
Of the fixed C, a higher proportion is consumed by respiration, in part because of higher leaf respiration ($R_{dark}$, Table 2) and in part because of higher sapwood volume (??confirm??) [**REFS**; @IncreasingRespiratoryCosts?].

>>>>

```{r eval=FALSE, echo=FALSE}
# refs in this table need to go in the nocite field below (easiest in alphabetical order)
```
---
nocite: |
@mau_temperate_2018; @cobleLightDrivesVertical2014; @sackHowStrongIntracanopy2006; @chinLeafAcclimationLight2017; @wykaResponsesLeafStructure2012; @athertonSpatialVariationLeaf2017; @kenzoHeightrelatedChangesLeaf2015; @kusiPlasticLeafMorphology2020; @dangProfilesPhotosyntheticallyActive1997;@gebauerEffectsProlongedDrought2015; @marencoLeafTraitPlasticity2017; @kafutiFoliarWoodTraits2020; @vanwittenbergheVariabilityStomatalConductance2012; @zhang_photosynthetic_2012; @weerasingheCanopyPositionAffects2014; @oldhamHydrostaticGradientNot2010; @ichieEcologicalDistributionLeaf2016; @gregoriouEffectsReducedIrradiance2007; @levizouNondestructiveAssessmentLeaf2005; @liakouraTrichomeDensityIts1997; @fauset_differences_2018; @niinemetsAcclimationHighIrradiance1998; Ishida et al. 1999; @millenLeafAngleAdaptive1979; @smithShootStructuralEffects1988; @hadleyInfluenceKrummholzMat1987; @panditharathnaChangeLeafStructure2008; @baltzerLeafOpticalResponses2005; @dietzVerticalPatternsDuration2007; @cobleHowVerticalPatterns2016a; @scartazzaInvestigatingEuropeanBeech2016; @duursmaVerticalCanopyGradients2006; @harleyEffectsLightTemperature1996; @hernandezSimilarTemperatureDependence2020; @turnbullScalingFoliarRespiration2003; @chenLeafEconomicsSpectrum2020; @vandewegPhotosyntheticParametersDark2012; @cavaleri_foliar_2008; @konigerXanthophyllCyclePigmentsPhotosynthetic1995;@matsubaraSunshadePatternsLeaf2009; @harrisChangesLeafProperties2013; @hansenVariationPigmentComposition2002a; @poorterLeafOpticalProperties1995; @cobleHowVerticalPatterns2016; @niinemetsCanopyGradientsLeaf2004; @poorterLeafOpticalProperties2000
---

@@

Further, higher heat tolerance in exposed canopies is facilitated by isoprenes in species capable of emitting isoprenes [*REF*].

@@

```{r eval=FALSE, echo=FALSE}
# refs in this table need to go in the nocite field below (easiest in alphabetical order)
```
---
nocite: | @kafutiFoliarWoodTraits2020; @vanwittenbergheVariabilityStomatalConductance2012a; @roberts_stomatal_1990; @dangProfilesPhotosyntheticallyActive1997; @marencoLeafTraitPlasticity2017; @ambroseEffectsHeightTreetop2010; @zweifelMiddayStomatalClosure2002; @slot_photosynthetic_2019; @hernandezSimilarTemperatureDependence2020; @urbanInductionPhotosynthesisImportance2007; @carterWithinCanopyExperimentalLeaf2018a; @martin_boundary_1999; @mau_temperate_2018; @kosugiSeasonalFluctuationsTemperature2006; @niinemetsVariabilityLeafMorphology2015; @bachofenLightVPDGradients2020; @hamerlynckPhotosyntheticStomatalResponses1996a; @cobleHowVerticalPatterns2016; @wykaResponsesLeafStructure2012; @rijkersEffectTreeHeight2000a; Ishida et al. 1999; @weerasingheCanopyPositionAffects2014; @scartazzaInvestigatingEuropeanBeech2016; @millerOnlySunlitLeaves2021; @harrisChangesLeafProperties2013; @legnerWithincanopyVariationPhotosynthetic2014; @kitaoCanopyNitrogenDistribution2018; @fauset_differences_2018; @rey-sanchez_spatial_2016; @mullerEvidenceEfficientNonevaporative2021; @curtis_intracanopy_2019; @meirLeafRespirationTwo2001; @turnbullScalingFoliarRespiration2003; @arakiVerticalSeasonalVariations2017; @bolstad_foliar_1999; @kenzoHeightrelatedChangesLeaf2015; @harleyEffectsLightTemperature1996;@xuSeasonalVariationTemperature2006; @harleyEnvironmentalControlsIsoprene1997; @niinemetsHowLightTemperature2015; @sharkeyFutureIsopreneEmission2014b; @sharkeyFutureIsopreneEmission2014; @simpragaVerticalCanopyGradient2013; @athertonSpatialVariationLeaf2017
---

@@

**Vertical gradients affect tree metabolism and growth.**
Whole-tree transpiration and gross primary production (GPP) increase in tandem with increasing tree size, where size is most commonly measured diameter breast height (DBH), but functionally linked to metrics such as foliage biomass, crown volume, or height, all of which scale with DBH [**REFS**, @anderson-teixeiraSizerelatedScalingTree2015].
Specifically, increases are linked increased leaf area and in part to increasing probability that the crown is in the canopy [@muller-landauTestingMetabolicEcology2006], where higher light availability and associated elevation of $g_s$ and $A_A$ result in higher leaf area-specific metabolism (Table 2).
Of the fixed C, a higher proportion is consumed by respiration, in part because of higher leaf respiration ($R_{dark}$, Table 2) and in part because of higher sapwood volume (??confirm??) [**REFS**; @IncreasingRespiratoryCosts?].
